Three 40 Million Year Old Mini-Neptunes from Kepler, TESS, and Gaia

### 2022-04-04 papers/ms.tex
update ms

% 208 words (250 max)
\begin{abstract}
  Stellar positions and velocities from Gaia are yielding a new and
  refined view of how stellar clusters evolve.
  Here we present an analysis of a group of $\approx$40 million
  year old stars spanning Cepheus ($l=100^\circ$) to Hercules
  ($l=40^\circ$), hereafter the Cep-Her complex.
  The group includes four known Kepler Objects of Interest:
  Kepler-1627 Ab ($R_{\rm p} = 3.85 \pm 0.11\,R_\oplus$, $P = 7.2\ {\rm days}$),
  Kepler-1643 b ($R_{\rm p} = 2.32 \pm 0.14\,R_\oplus$, $P = 5.3\ {\rm days}$),
  KOI-7368 b ($R_{\rm p} = 2.22 \pm 0.12\,R_\oplus$, $P = 6.8\ {\rm days}$), and
  KOI-7913 Ab ($R_{\rm p} = 2.34 \pm 0.18\,R_\oplus$, $P = 24.2\ {\rm days}$).
  Kepler-1627 is a Neptune-sized planet in a sub-component of the
  complex called the $\delta$\ Lyr\ cluster
  \citep{bouma_kep1627_2022}.
  Here we focus on the latter three systems, which are in other
  sub-components of the complex (RSG-5 and CH-2).
  Based on kinematic evidence from Gaia, stellar rotation periods from
  TESS, and spectroscopy, these three systems are also $\approx$40
  million years old.
  More specifically, we find that Kepler-1643 (in RSG-5) is
  $46^{+9}_{-7}$\,Myr old, while KOI-7368 and KOI-7913 (in CH-2) are
  $36^{+10}_{-8}$\,Myr old.
  Based on the transit shapes and high resolution imaging, they are
  all most likely planets, with false positive probabilities of
  $6\times10^{-9}$, $5\times10^{-3}$, and $1\times10^{-4}$ for
  Kepler-1643, KOI-7368, and KOI-7913 respectively.
  These planets are therefore the first empirical
  demonstration that mini-Neptunes with sizes of $\approx$2 Earth
  radii exist at ages of $\approx$40 million years.
\end{abstract}

@@

%(e.g., Jenkins KSCI-19081-002 KPDH pp 129, or many other pages)
KOI-7913 appears to be a binary.\footnote{Relevant identifiers for the
primary are KIC 8873450 = Gaia DR2/EDR3 2106235301785454208.  For the
secondary, they are KIC 8873448 = Gaia DR2/EDR3 2106235301785453824.}
The north-west primary is $\approx$0.5 magnitudes brighter than the
south-east secondary in optical passbands.  The two stars are
separated in Gaia EDR3 by $3\farcs54$ on-sky, and have parallaxes
consistent within $1$-$\sigma$ (with an average $\varpi=3.66 \pm
0.01$\,mas).  This corresponds to an apparent on-sky separation of
$959 \pm 2$ au.  Gaia EDR3 quotes
$(\mu_{\alpha'},\mu_\delta)=(2.439\pm0.015, 1.009\pm 0.016)\,{\rm
mas}\,{\rm yr}^{-1}$ for the primary, and
$(\mu_{\alpha'},\mu_\delta)=(2.696\pm0.020, 0.650 \pm 0.021) \,{\rm
mas}\,{\rm yr}^{-1}$ for the secondary.  Since two stars were resolved
in the Kepler Input Catalog and are roughly one Kepler pixel apart, an
accurate crowding metric has already been applied in the NASA Ames
data products to correct the mean flux level
\citep{2017ksci.rept....6M}.  This is important for deriving accurate
transit depths.

@@

\subsection{Transit and Stellar Variability Model}

@@

We fitted the models using \texttt{exoplanet}
\citep{exoplanet:exoplanet}.  We assumed a Gaussian likelihood, and
sampled using \texttt{PyMC3}'s No-U-Turn Sampler
\citep{hoffman_no-u-turn_2014}, after having initialized using the
maximum {\it a posteriori} (MAP) model.  We used the
\citet{gelman_inference_1992} statistic ($\hat{R}$) as our convergence
diagnostic.  Abbreviated results are in Table~\ref{fig:sysparams}, the
full results are in Appendix~\ref{app:transit}.

One weakness of our fitting approach is that we have fixed 5 implicit
parameters per individual transit to their MAP values to remove the
stellar variability signal.  An alternative could be to fit the
planetary transits simultaneously with the starspot-induced
variability using, say, a quasiperiodic Gaussian process (GP).  We
explored this approach, but found that it required careful fine-tuning
of the hyperparameter priors, otherwise the GP would tend to
incorporate anything that was not part of the transit, even if it was
short-term variability attributable to for instance flares, or simply
instrumental noise.  This failure mode is particularly pernicious, in
that it yields an inflated and ill-founded sense of confidence in the
transit model results.  While our model is simpler, it has the benefit
that its white noise jitter never trades off with any parameter
equivalent to a damping timescale for the coherence of the GP.  It is
also computationally efficient, while capturing the planetary
parameters that we actually care about.

@@

@@

\software{
  %\texttt{arviz} \citep{arviz_2019},
  %\texttt{altaipony} \citep{ilin_flares_2021},
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
  %\texttt{astroplan} \citep{astroplan2018},
	%\texttt{AstroImageJ} \citep{collins_astroimagej_2017},
  \texttt{astropy} \citep{astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  %\texttt{BATMAN} \citep{kreidberg_batman_2015},
  %\texttt{ceres} \citep{brahm_2017_ceres},
  %\texttt{cdips-pipeline} \citep{bhatti_cdips-pipeline_2019},
  \texttt{corner} \citep{corner_2016},
  %\texttt{emcee} \citep{foreman-mackey_emcee_2013},
  \texttt{exoplanet} \citep{exoplanet:exoplanet}, and its
  dependencies \citep{exoplanet:agol20, exoplanet:kipping13, exoplanet:luger18,
   	exoplanet:theano},
	%\texttt{gala} \citep{gala,PriceWhelan_2017_gala_zenodo},
	%\texttt{IDL Astronomy User's Library} \citep{landsman_1995},
  %\texttt{IPython} \citep{perez_2007},
	%\texttt{isochrones} \citep{morton_2015_isochrones},
	%\texttt{lightkurve} \citep{lightkurve_2018},
  %\texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  %\texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
  %\texttt{numpy} \citep{walt_numpy_2011}, 
  %\texttt{pandas} \citep{mckinney-proc-scipy-2010},
  %\texttt{pyGAM} \citep{serven_pygam_2018_1476122},
  \texttt{PyMC3} \citep{salvatier_2016_PyMC3},
  %\texttt{radvel} \citep{fulton_radvel_2018},
  %\texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{jones_scipy_2001},
  %\texttt{TESS-point}  \citep{burke_2020},
  %\texttt{tesscut} \citep{brasseur_astrocut_2019},
	%\texttt{VESPA} \citep{morton_efficient_2012,vespa_2015},
  %\texttt{webplotdigitzer} \citep{rohatgi_2019},
  %\texttt{wotan} \citep{hippke_wotan_2019}.
}
\

@@

Table goes here

>>>>

% 208 words (250 max)
\begin{abstract}
  Stellar positions and velocities from Gaia are yielding a new and
  refined view of how stellar clusters evolve.
  Here we present an analysis of a group of $\approx$40 million
  year old stars spanning Cepheus ($l=100^\circ$) to Hercules
  ($l=40^\circ$), hereafter the Cep-Her complex.
  The group includes four known Kepler Objects of Interest:
  Kepler-1627 Ab ($R_{\rm p} = 3.85 \pm 0.11\,R_\oplus$, $P = 7.2\ {\rm days}$),
  Kepler-1643 b ($R_{\rm p} = 2.32 \pm 0.14\,R_\oplus$, $P = 5.3\ {\rm days}$),
  KOI-7368 b ($R_{\rm p} = 2.22 \pm 0.12\,R_\oplus$, $P = 6.8\ {\rm days}$), and
  KOI-7913 Ab ($R_{\rm p} = 2.34 \pm 0.18\,R_\oplus$, $P = 24.2\ {\rm days}$).
  Kepler-1627 is a Neptune-sized planet in a sub-component of the
  complex called the $\delta$\ Lyr\ cluster
  \citep{bouma_kep1627_2022}.
  Here we focus on the latter three systems, which are in other
  sub-components of the complex (RSG-5 and CH-2).
  Based on kinematic evidence from Gaia, stellar rotation periods from
  TESS, and spectroscopy, these three systems are also $\approx$40
  million years old.
  More specifically, we find that Kepler-1643 (in RSG-5) is
  $46^{+9}_{-7}$\,Myr old, while KOI-7368 and KOI-7913 (in CH-2) are
  $36^{+10}_{-8}$\,Myr old.
  Based on the transit shapes and high resolution imaging, they are
  all most likely planets, with false positive probabilities of
  $6\times10^{-9}$, $4\times10^{-3}$, and $1\times10^{-4}$ for
  Kepler-1643, KOI-7368, and KOI-7913 respectively.
  These planets are therefore the first empirical
  demonstration that mini-Neptunes with sizes of $\approx$2 Earth
  radii exist at ages of $\approx$40 million years.
\end{abstract}

@@

%(e.g., Jenkins KSCI-19081-002 KPDH pp 129, or many other pages)
KOI-7913 is almost certainly a binary.\footnote{Relevant identifiers for the
primary are KIC 8873450 = Gaia DR2/EDR3 2106235301785454208.  For the
secondary, they are KIC 8873448 = Gaia DR2/EDR3 2106235301785453824.}
The north-west primary is $\approx$0.5 magnitudes brighter than the
south-east secondary in optical passbands.  The two stars are
separated in Gaia EDR3 by $3\farcs54$ on-sky, and have parallaxes
consistent within $1$-$\sigma$ (with an average $\varpi=3.66 \pm
0.01$\,mas).  This corresponds to an apparent on-sky separation of
$959 \pm 2$ au.  Gaia EDR3 quotes
$(\mu_{\alpha'},\mu_\delta)=(2.439\pm0.015, 1.009\pm 0.016)\,{\rm
mas}\,{\rm yr}^{-1}$ for the primary, and
$(\mu_{\alpha'},\mu_\delta)=(2.696\pm0.020, 0.650 \pm 0.021) \,{\rm
mas}\,{\rm yr}^{-1}$ for the secondary.  Since two stars were resolved
in the Kepler Input Catalog and are roughly one Kepler pixel apart, an
accurate crowding metric has already been applied in the NASA Ames
data products to correct the mean flux level
\citep{2017ksci.rept....6M}.  This is important for deriving accurate
transit depths.

@@

\subsection{Transit and Stellar Variability Model}
\label{sec:fitting}

@@

We fitted the models using \texttt{exoplanet}
\citep{exoplanet:exoplanet}.  We assumed a Gaussian likelihood, and
sampled using \texttt{PyMC3}'s No-U-Turn Sampler
\citep{hoffman_no-u-turn_2014}, after having initialized using the
maximum {\it a posteriori} (MAP) model.  We used the
\citet{gelman_inference_1992} statistic ($\hat{R}$) as our convergence
diagnostic.  The results are shown in the lower panels of
Figure~\ref{fig:planets}, and the important derived parameters are in
Table~\ref{tab:sysparams}.  The set of full parameters with their
priors are given in Appendix~\ref{app:transit}.

One weakness of our fitting approach is that we have fixed 5 implicit
parameters per individual transit to their MAP values to remove the
stellar variability signal.  An alternative could be to fit the
planetary transits simultaneously with the starspot-induced
variability using, say, a quasiperiodic Gaussian process (GP).  We
explored this approach, but found that it required careful fine-tuning
of the hyperparameter priors, otherwise the GP would tend to
incorporate any variability that was not part of the transit, even if it was
short-term variability attributable to flares, or simply
instrumental noise.  This failure mode is somewhat pernicious, in that
it yields an ill-founded sense of confidence in an ``overly clean''
transit model fit.  Although our model is simpler, it has the benefit
that the white noise jitter never trades off with any parameter
equivalent to a damping timescale for the coherence of the GP.  It is
also computationally efficient, and it captures the planetary
parameters that we actually care about.

@@

In the future, it may be possible to obtain independent evidence for
the Cep-Her planets that directly confirms their planetary nature.
The most likely path for that would be through transit
observations with a high-resolution high-throughput spectrograph
\citep[see][]{bouma_kep1627_2022}.  In the interim, it is of some
interest whether any type of astrophysical false positive is likely to
explain the transit signals, or whether they can be argued to be
planetary on a statistical basis.  We adopt the Bayesian framework
implemented in \texttt{VESPA} for this purpose
\citep{morton_efficient_2012,vespa_2015}.  Briefly summarized, the
priors in \texttt{VESPA} take as given the binary star occurrence rate
from \citet{raghavan_survey_2010}, direction-specific star counts from
\citet{girardi_star_2005}, and planet occurrence rates as described by
\citet[][Section~3.4]{morton_efficient_2012}.  The likelihoods are
then evaluated by forward-modeling a synthetic population of eclipsing
bodies for each possible astrophysical model class, in which each
population member has a particular trapezoidal eclipse depth, total
duration, and ingress duration.  These summary statistics are then
compared against the actual photometric data to evaluate the 
probabilities of false positive scenarios such as foreground eclipsing
binaries, hierarchical eclipsing binaries, and background eclipsing
binaries.

\paragraph{Kepler-1643}
Kepler-1643~b (KOI-6186.01) was already validated as a transiting
planet by \citet{morton_false_2016}.  Based on the Q1-Q17 DR25 False
Positive Probability table hosted at the NASA Exoplanet Archive, they
found a probability for any of the considered false positive scenarios
of 9$\times$10$^{-6}$.  Repeating the calculation with our own
stellar-variability correction and the new NIRC2 imaging constraints,
we find ${\rm FPP} = 6\times10^{-9}$, where the improvement is
presumably due to the new NIRC2 contrast curves.
Figure~\ref{fig:planets} shows the intuitive justification for the
strong statistical result: the transit is flat and has a reasonably
high S/N ($\approx 47$), which is difficult to reproduce with
eclipsing binary models.

An interesting note on Kepler-1643 is that it formally fails the
default centroid shift test in both the \texttt{q1\_q17\_dr25\_koi}
and \texttt{q1\_q17\_dr25\_sup\_koi} releases, with the angular
distance between the target star position and the location of the
transiting source being reported as $0\farcs99$ (4.4-$\sigma$).  This
seems to be mostly due to two outlying quarters (Q2 and Q6) driving
the offset -- the centroid locations from the other Kepler quarters
are all consistent within $\lesssim 0\farcs5$.  Nonetheless, this is
an instructive exercise in how highly variable stars will complicate
centroid-based vetting tests.  The centroid shifts measured by these
tests are determined from the in- and out-of-transit flux-weighted
centroids, but for our stars there is no static baseline in either in-
or out-of-transit phase.

\paragraph{KOI-7368}
KOI-7368.01 is listed on the NASA Exoplanet Archive as a ``CANDIDATE''
planet.  The Q1-Q17 DR25 False Positive Probability table however
lists its probability of being an astrophysical false positive as
``1.0''. However, given that none of the other columns in this table
are populated, this is most likely a database entry error ({\bf ask
Jessie}).  Performing the validation calculation using our new NIRC2
imaging constraints, we find ${\rm FPP} \approx 4\times10^{-3}$.
Though not as convincing as Kepler-1643, this formally clears the
usual threshold for calling the planet statistically validated
\citep{morton_efficient_2012}.  The S/N of the transit itself is
$\approx 32$, which indicates that it is unlikely to be caused by
systematic noise in the light curve.  Visually, one can also inspect
each individual transit window in the light curve itself.  This
exercise shows a weak notch in each case; over the 198 observed
transits, the resulting transit signal seems sensible.

Why was KOI-7368 not previously validated?  One reason could be the
increased stellar variability, or the lower S/N of the transit.
Another reason could be that it also formally fails the centroid shift
test, just like Kepler-1643.  In this case, the reported offset is
somewhat less significant ($0\farcs22$; 3.0-$\sigma$).  Again,
inspection of the data validation reports shows that this is caused by
a few outlying quarters: in this case Q4, Q5, Q8, and Q12.
The fact that the remaining quarters show a far more consistent
scatter suggests that these outliers are caused by local stellar
variability, rather than an astrophysical offset.
Our imaging also shows that there are no neighboring sources that
could cause the offset.

\paragraph{KOI-7913}
KOI-7913.01 is listed on the NASA Exoplanet Archive as a ``CANDIDATE''
planet.  The Q1-Q17 DR25 False Positive Probability table lists its
probability of being an astrophysical false positive as
$1.4\times10^{-4}$, with the most likely false positive scenario being
a blended eclipsing binary.  Repeating the calculation with our new
detrending and contrast curves, we find a similar result: ${\rm FPP} =
1.3\times10^{-4}$.  While the transit has the lowest S/N of any of the
objects discusssed (${\rm S/N}=14$), its marginally lower FPP relative
to KOI-7368 can be understood through its relatively flat-bottomed
shape, combined with its long transit duration relative to plausible
eclipsing binary models.

The disposition of KOI-7913.01 seems to have been previously debated:
in \texttt{q1\_q17\_dr25\_koi} it was flagged as a ``FALSE POSITIVE'',
with the comment ``CENT\_KIC\_POS---HALO\_GHOST''.  This comment and
disposition seem to have been removed in the
\texttt{q1\_q17\_dr25\_sup\_koi} data release, which renamed it a
``CANDIDATE'', without any flag given.  As described by
\citet{thompson_planetary_2018}, the ``CENT\_KIC\_POS'' flag is an
indication that the measured source centroid is offset from its
expected location in the Kepler Input Catalog.  The final Data
Validation Reports (generated 2016 Jan 30) do not seem to show this,
and the statistical significance of any centroid offset seems to be
lower than for KOI-7368 and Kepler-1643 (which, as we have argued,
both show centroid offsets that seem to be explained by the stellar
variability).

What of the ``HALO\_GHOST'' flag?  This test measures the transit
strength for the pixels inside the aperture, and the ring of pixels
around that aperture (the ``halo'').  Generally one expects the
transit signal to be strongest in the central aperture.  Two types
of false positive scenarios can change this, triggering the flag: the first is
when optical ghosts from bright eclipsing binaries reflect off the
CCD, and contaminate the target star.  The second is when the PRF of
nearby stars directly overlaps with the PRF of the target star (see
\citealt{thompson_planetary_2018}, Section A.5.2).

The most obvious explanation for KOI-7913 is the latter case, since
KOI-7913 B is $\approx0.9$ Kepler pixels away from Kepler-7913 A.
Due to the on-sky orientation of KOI-7913 A and
KOI-7913 B, the default ``optimal
aperture'' selected in Q3, Q7, Q11, and Q15 in fact included both
stars, while for the remaining quarters KOI-7913 B was (correctly)
excluded from the optimal aperture
(see pages 35 through 71 of the Data Validation Report.)

{\bf
TODO: words!  Describe Halo ghost statistic.  Why should I, or
shouldn't I, be worried?
}

@@

\software{
  %\texttt{arviz} \citep{arviz_2019},
  %\texttt{altaipony} \citep{ilin_flares_2021},
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
  %\texttt{astroplan} \citep{astroplan2018},
	%\texttt{AstroImageJ} \citep{collins_astroimagej_2017},
  \texttt{astropy} \citep{astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  %\texttt{BATMAN} \citep{kreidberg_batman_2015},
  %\texttt{ceres} \citep{brahm_2017_ceres},
  %\texttt{cdips-pipeline} \citep{bhatti_cdips-pipeline_2019},
  %\texttt{corner} \citep{corner_2016},
  %\texttt{emcee} \citep{foreman-mackey_emcee_2013},
  \texttt{exoplanet} \citep{exoplanet:exoplanet}, and its
  dependencies \citep{exoplanet:agol20, exoplanet:kipping13, exoplanet:luger18,
   	exoplanet:theano},
	%\texttt{gala} \citep{gala,PriceWhelan_2017_gala_zenodo},
	%\texttt{IDL Astronomy User's Library} \citep{landsman_1995},
  %\texttt{IPython} \citep{perez_2007},
	%\texttt{isochrones} \citep{morton_2015_isochrones},
	%\texttt{lightkurve} \citep{lightkurve_2018},
  %\texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  %\texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
  %\texttt{numpy} \citep{walt_numpy_2011}, 
  %\texttt{pandas} \citep{mckinney-proc-scipy-2010},
  %\texttt{pyGAM} \citep{serven_pygam_2018_1476122},
  \texttt{PyMC3} \citep{salvatier_2016_PyMC3},
  %\texttt{radvel} \citep{fulton_radvel_2018},
  %\texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{jones_scipy_2001},
  %\texttt{TESS-point}  \citep{burke_2020},
  %\texttt{tesscut} \citep{brasseur_astrocut_2019},
	\texttt{VESPA} \citep{morton_efficient_2012,vespa_2015},
  %\texttt{webplotdigitzer} \citep{rohatgi_2019},
  %\texttt{wotan} \citep{hippke_wotan_2019}.
}
\

@@

Tables~\ref{tab:kepler1643}, \ref{tab:koi7368}, \ref{tab:koi7913} give
the full set of fitted and derived parameters from the model described in
Section~\ref{sec:fitting}.
Priors and convergence statistics are also listed.

\input{source_Kepler_1643_simpletransit_posteriortable.tex}
\input{source_KOI_7368_simpletransit_posteriortable.tex}
\input{source_KOI_7913_simpletransit_posteriortable.tex}
